---
title: "Synthetic Data Generation for Economic Policymaking"
author:
- Austin Denteh
- Prince Baah
- Blaise Bayuo
date: "June 24, 2026"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  pdf_document:
    toc: true
subtitle: "GLSS6 2012/13 -> DHS 2014"
params:
  subsample_train: 0
  subsample_holdout: 0
  subsample_dhs: 0
  m_implicates: 5
  maxit: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 4.8
)

library(dplyr)
library(ggplot2)
library(mice)

source("code/utils/project_paths.R")

# ---- SCALE KNOBS (set in the YAML params above) --------------------------
# subsample_train / subsample_holdout / subsample_dhs: 0 = use ALL rows;
#   a positive number subsamples that many rows (fixed seed) to speed up the
#   live mice fit. m_implicates / maxit control the number of imputed datasets and
#   chained-equation iterations. mice cycles because each conditional model uses
#   current draws of the other incomplete variables; repeated cycles let the
#   welfare and NHIS draws settle into a mutually consistent imputed data set.
#   Full real-data run: 0 / 0 / 0 / 5 / 5.
#   Fast classroom demo, e.g.: 3000 / 1000 / 3000 / 3 / 2.
# --------------------------------------------------------------------------
DATA_FAMILY <- "teaching"
# Source: GLSS6 poverty/welfare aggregate file
# AGGREGATES/POV_GHA_2013_GLSS6-updated.dta, variables
# poveline_fd_2013 (extreme poverty line per year) and
# poveline_2013 (absolute poverty line per year).
POV_EXTREME <- 792.05
POV_ABSOLUTE <- 1314
SYNTH_SEED <- 20260614
EXTENSION_SEED <- 20260616
M <- params$m_implicates
MAXIT <- params$maxit

input_dir <- project_path("data", "practical", "teaching")
input_prefix <- "teach_"

diag_dir <- project_path("output", "diagnostics")

need_file <- function(path) {
  if (!file.exists(path)) stop("Missing required file: ", path, call. = FALSE)
  path
}

read_input <- function(name) {
  readRDS(need_file(file.path(input_dir, paste0(input_prefix, name, ".rds"))))
}

# ---- WEIGHTING HELPERS ---------------------------------------------------
# Every rate we REPORT is survey-weighted. (Slide "Decisions for every imputed
# variable": the model is fit unweighted, but summaries and
# all validation statistics carry household_weight.) These helpers centralise
# that rule so no chunk forgets the weights.

# Weighted mean that drops NA in either x or the weight.
w_mean <- function(x, w) {
  ok <- !is.na(x) & !is.na(w)
  if (!any(ok)) return(NA_real_)
  weighted.mean(x[ok], w[ok])
}

# Weighted mean of a 0/1 indicator, expressed as a percentage.
w_rate <- function(x, w) 100 * w_mean(as.numeric(x), w)

# Weighted category shares for one variable, tidy long format (for side-by-side
# GLSS vs DHS comparisons in the warm-up).
w_tab <- function(dat, var, weight = "household_weight", source_label = NULL) {
  d <- dat %>%
    filter(!is.na(.data[[var]]), !is.na(.data[[weight]])) %>%
    mutate(value = as.character(.data[[var]]))
  d %>%
    group_by(value) %>%
    summarise(n = n(),
              pct_w = 100 * sum(.data[[weight]]) / sum(d[[weight]]),
              .groups = "drop") %>%
    mutate(variable = var, survey = source_label, pct_w = round(pct_w, 1)) %>%
    select(survey, variable, value, pct_w, n)
}

# ---- RUBIN'S RULES POOLING -----------------------------------------------
# Combine an estimate computed on each of the m imputed datasets into one number with
# correct uncertainty. (Slide "Multiple imputation keeps uncertainty visible"
# and backup slide "Combining estimates across imputed datasets".)
#   qbar  = average of the m point estimates
#   ubar  = average WITHIN-imputation variance (uncertainty if welfare were known)
#   b     = BETWEEN-imputation variance        (extra uncertainty from imputing it)
#   total = ubar + (1 + 1/m) * b               (Rubin's total variance)
pool_scalar <- function(q, u) {
  m <- length(q)
  qbar <- mean(q, na.rm = TRUE); ubar <- mean(u, na.rm = TRUE)
  b <- stats::var(q, na.rm = TRUE)
  total <- ubar + (1 + 1 / m) * b
  data.frame(estimate = qbar, within_var = ubar, between_var = b,
             total_var = total, se = sqrt(total))
}

# Weighted proportion plus its within-imputation variance, using an effective
# sample size (Kish) for the weights. Feeds the WITHIN term of pool_scalar().
weighted_binary_est <- function(x, w) {
  ok <- !is.na(x) & !is.na(w)
  x <- as.numeric(x[ok]); w <- w[ok]
  p <- weighted.mean(x, w)
  n_eff <- sum(w)^2 / sum(w^2)
  data.frame(estimate = p, var = p * (1 - p) / n_eff)
}

# Recovery metrics for the holdout: how close are imputed (log) welfare draws to
# the TRUE held-out welfare? Used by the engine-switch extension.
recovery_metrics <- function(dat, label) {
  dat %>%
    group_by(.imp) %>%
    summarise(
      mae_log = mean(abs(synth_log_welfare - log(true_welfare)), na.rm = TRUE),
      rmse_log = sqrt(mean((synth_log_welfare - log(true_welfare))^2, na.rm = TRUE)),
      corr_log = cor(synth_log_welfare, log(true_welfare), use = "complete.obs"),
      synth_poverty_rate_w = w_rate(synth_poor, household_weight),
      true_poverty_rate_w = w_rate(true_poor, household_weight),
      .groups = "drop"
    ) %>%
    summarise(engine = label, mae_log = mean(mae_log), rmse_log = mean(rmse_log),
              corr_log = mean(corr_log),
              synth_poverty_rate_w = mean(synth_poverty_rate_w),
              true_poverty_rate_w = mean(true_poverty_rate_w), .groups = "drop")
}
```

This practical shows how to enrich DHS 2014 with GLSS6 welfare aggregate variables.
The idea follows the theory deck step by step:

1. Learn the relationship between welfare-like targets and shared predictors in GLSS6
   (slide *"Synthetic data is a missing data framework"*: we estimate $p(Y \mid X)$ in GLSS).
2. Check whether the learned model recovers hidden GLSS holdout welfare
   (slide *"Validation Check 1 — GLSS holdout"*).
3. Run one transport check in DHS using an observed target
   (slide *"Validation Check 2 — transport"*).
4. **Impute** multiple welfare datasets for DHS households
   (slides *"The imputation workflow at a glance"* and *"Multiple imputation keeps uncertainty visible"*).
5. Analyze DHS child outcomes by imputed poverty status
   (slide *"Running question for today"*).

This is the **live-compute** version: the model is fitted inside this notebook. Use the **scale knobs** in the YAML `params` to subsample rows or reduce imputed datasets while
you develop, then run the full settings for the real result. The notebook uses prepared teaching data files. 

> **Vocabulary note.** Throughout, we say we **impute** (fill in) the missing welfare block, or
> **predict** it when we transport the fitted model to new households. The literature calls the
> filled-in values "synthetic," so code columns prefixed `synth_` simply hold these **imputed**
> draws. The DHS transport check (Section *"DHS Transport Check: How well does imputed electricity transfer to DHS"*) is what that literature calls
> *calibration*.

# Dataset Warm-Up

We start by loading the prepared household and child files.

```{r load-data}
# Four prepared teaching files = the four corners of the design (slide
# "Our two data sources: GLSS and DHS"):
#   glss_train     GLSS6 rows that TRAIN the model  (X and welfare Y both observed)
#   glss_holdout   GLSS6 rows held out for Validation Check 1 (welfare known, but we hide it)
#   dhs_households the records to be ENRICHED        (X only; welfare missing by design)
#   dhs_children  child outcomes linked to DHS households (the analysis outcomes)
glss_train <- read_input("glss6_train")
glss_holdout <- read_input("glss6_holdout")
dhs_households <- read_input("dhs2014_households")
dhs_children <- read_input("dhs2014_children")

# Apply the scale knobs (0 = use all rows). Children follow the DHS subsample.
set.seed(SYNTH_SEED)
if (params$subsample_train  > 0 && params$subsample_train  < nrow(glss_train))
  glss_train    <- glss_train    %>% slice_sample(n = params$subsample_train)
if (params$subsample_holdout > 0 && params$subsample_holdout < nrow(glss_holdout))
  glss_holdout  <- glss_holdout  %>% slice_sample(n = params$subsample_holdout)
if (params$subsample_dhs    > 0 && params$subsample_dhs    < nrow(dhs_households))
  dhs_households <- dhs_households %>% slice_sample(n = params$subsample_dhs)
dhs_children <- dhs_children %>% filter(hh_key %in% dhs_households$hh_key)

surface_counts <- data.frame(
  file = c("GLSS train households", "GLSS holdout households",
           "DHS target households", "DHS children"),
  rows = c(nrow(glss_train), nrow(glss_holdout),
           nrow(dhs_households), nrow(dhs_children)),
  columns = c(ncol(glss_train), ncol(glss_holdout),
              ncol(dhs_households), ncol(dhs_children))
)

knitr::kable(surface_counts)
```

## What Each Survey Contributes

GLSS is the source survey for welfare. DHS is the target survey for child health and nutrition outcomes. The bridge between them is the set of shared household predictors.

```{r observed-missing-map}
# This is the "missing-data framework" slide in table form: each concept is
# Observed or Missing in each file. The whole exercise is filling the
# "Missing by design" cells in DHS using the GLSS-learned model.
observed_map <- data.frame(
  concept = c("Shared household predictors",
              "Welfare aggregate",
              "Derived poverty status",
              "Any household member registered with NHIS",
              "DHS wealth benchmark",
              "Child nutrition and illness outcomes"),
  glss = c("Observed", "Observed", "Observed from welfare",
           "Observed", "Not observed", "Not observed"),
  dhs_households = c("Observed", "Missing by design",
                     "Missing by design", "Observed (transport target)",
                     "Observed", "Household link only"),
  dhs_children = c("Linked through household", "Not observed",
                   "Not observed", "Linked through household",
                   "Linked through household", "Observed")
)

knitr::kable(observed_map)
```

Now look at a few shared predictors. These are not identical surveys, so the goal is not perfect balance.

```{r shared-predictor-summary}
binary_summary <- data.frame(
  survey = c("GLSS train", "DHS households"),
  urban_pct_w = c(
    w_rate(glss_train$urban, glss_train$household_weight),
    w_rate(dhs_households$urban, dhs_households$household_weight)
  ),
  electricity_pct_w = c(
    w_rate(glss_train$has_electricity, glss_train$household_weight),
    w_rate(dhs_households$has_electricity, dhs_households$household_weight)
  ),
  female_head_pct_w = c(
    w_rate(glss_train$head_female, glss_train$household_weight),
    w_rate(dhs_households$head_female, dhs_households$household_weight)
  ),
  mean_hh_size_w = c(
    w_mean(glss_train$hh_size, glss_train$household_weight),
    w_mean(dhs_households$hh_size, dhs_households$household_weight)
  )
) %>%
  mutate(across(where(is.numeric), ~ round(.x, 1)))

knitr::kable(binary_summary)
```

```{r region-side-by-side}
region_summary <- bind_rows(
  w_tab(glss_train, "region10", source_label = "GLSS train"),
  w_tab(dhs_households, "region10", source_label = "DHS households")
) %>%
  select(variable, value, survey, pct_w) %>%
  tidyr::pivot_wider(names_from = survey, values_from = pct_w)

knitr::kable(region_summary)
```

```{r material-side-by-side}
housing_summary <- bind_rows(
  w_tab(glss_train, "floor_material_screen", source_label = "GLSS train"),
  w_tab(dhs_households, "floor_material_screen", source_label = "DHS households"),
  w_tab(glss_train, "water_source_cat", source_label = "GLSS train"),
  w_tab(dhs_households, "water_source_cat", source_label = "DHS households")
) %>%
  select(variable, value, survey, pct_w) %>%
  tidyr::pivot_wider(names_from = survey, values_from = pct_w) %>%
  arrange(variable, value)

knitr::kable(housing_summary)
```

Before modeling, look directly at the two variables in the synthetic target block: the welfare aggregate and household NHIS registration. In the notebook, we synthesize
\[
Y = \{\log(\text{welfare}),\ \text{NHIS}\}
\]
for the rows where that block is masked or missing; poverty is then derived from each welfare draw. In probability-model language, the target is $p(Y \mid X)$, where $X$ is the shared predictor set. The `mice` implementation approximates that joint target with chained conditionals:
\[
p(\log(\text{welfare}),\ \text{NHIS} \mid X)
\quad \text{via} \quad
p(\log(\text{welfare}) \mid X,\ \text{NHIS})
\quad \text{and} \quad
p(\text{NHIS} \mid X,\ \log(\text{welfare})).
\]
The welfare distribution is skewed, which motivates the log-welfare model used later.

```{r warmup-welfare-nhis}
# First show welfare on its raw scale: the long right tail is the reason the
# imputation model works on log welfare.
ggplot(glss_train, aes(x = welfare)) +
  geom_histogram(bins = 45, fill = "#2f6f73", color = "white") +
  labs(x = "GLSS welfare, raw annual GHS per adult equivalent",
       y = "Households") +
  theme_minimal(base_size = 12)

# Then show the same variable on a log axis. This is why the model below
# imputes LOG welfare with a normal model and back-transforms with exp()
# (slides "Begin with single imputation" and "The imputation workflow at a glance").
ggplot(glss_train, aes(x = welfare)) +
  geom_histogram(bins = 45, fill = "#2f6f73", color = "white") +
  scale_x_log10() +
  labs(x = "GLSS welfare, log scale", y = "Households") +
  theme_minimal(base_size = 12)

nhis_warmup <- data.frame(
  survey = c("GLSS train", "DHS households"),
  nhis_any_member_registered_pct_w = c(
    w_rate(glss_train$nhis_any_member_registered, glss_train$household_weight),
    w_rate(dhs_households$nhis_any_member_registered,
           dhs_households$household_weight)
  )
) %>%
  mutate(nhis_any_member_registered_pct_w =
           round(nhis_any_member_registered_pct_w, 1))

knitr::kable(nhis_warmup)
```

## Weighted vs. Unweighted Summaries

The imputation model below is fit unweighted, because the primary `mice` engines used here do not accept survey weights. Summaries and validation statistics do use `household_weight`.

```{r weighted-unweighted-demo}
weighted_unweighted <- data.frame(
  statistic = c("GLSS poverty rate", "GLSS urban share"),
  unweighted = c(
    mean(glss_train$welfare < POV_ABSOLUTE, na.rm = TRUE),
    mean(glss_train$urban, na.rm = TRUE)
  ),
  weighted = c(
    w_mean(glss_train$welfare < POV_ABSOLUTE, glss_train$household_weight),
    w_mean(glss_train$urban, glss_train$household_weight)
  )
) %>%
  mutate(across(c(unweighted, weighted), ~ round(100 * .x, 1)))

knitr::kable(weighted_unweighted)
```

One working assumption is that, after conditioning on region, urban residence, and the other shared predictors, the unweighted fitting step is a reasonable teaching approximation.

# Run the Imputation (computed live)

Everything below is computed **inside this notebook**. The chunk fits `mice` on the GLSS training rows only (`ignore =`), draws `m` imputed datasets, derives poverty from each drawn welfare, and builds every object the later sections display (the holdout check, the
transport check, and the DHS enrichment). Reduce the scale knobs in `params` to make this run
faster while developing.

This single chunk is the engine room. Read it as four helper functions followed by three runs:

- `derive_poverty()` / `build_frame()` / `run_mice()` / `collect_draws()` — reusable machinery.
- **Run A** = the DHS enrichment, **Run B** = the GLSS holdout (Validation Check 1), **Run C** = the DHS
  electricity transport check (Validation Check 2). All three reuse the same machinery on different rows.

```{r run-imputation}
# --- Which predictors X bridge the two surveys? --------------------------
# The manifest is the variable-role table from data prep (slide "Activity:
# sort variables before modeling"): predictor_core = the shared predictors X.
manifest <- read.csv(need_file(file.path(diag_dir, "practical_variable_manifest.csv")),
                     stringsAsFactors = FALSE)
core_pred <- manifest$variable[manifest$manifest_role == "predictor_core"]

# Drop three columns from the MODEL only (kept in the extract for inspection):
# they are exact linear combinations of other predictors, which makes the
# regression unstable. members_total ~ hh_size; adults_15plus = members_total -
# children_under15; floor_non_earth = duplicate of the floor_material_screen dummy.
collinear_drop <- c("members_total", "adults_15plus", "floor_non_earth")
model_pred <- setdiff(core_pred, collinear_drop)

# cat_predictors is a helper list of categorical predictors that should be converted to factors before mice runs
cat_predictors <- intersect(
  c("region10", "water_source_cat", "cooking_fuel_cat",
    "roof_material_cat", "floor_material_screen"), core_pred)

# DERIVE, don't draw (slide "From the welfare aggregate to poverty").
# Poverty is a deterministic cut of welfare at Ghana's official lines, applied
# AFTER each welfare draw — never modelled as its own outcome.
derive_poverty <- function(welfare) {
  cut(welfare, c(-Inf, POV_EXTREME, POV_ABSOLUTE, Inf),
      labels = c("Very poor", "Poor", "Non poor"))
}


# build_frame prepares the data that mice will see.It takes an “apply” dataset, 
# either the GLSS holdout or DHS households, and stacks it underneath glss_train.
# It also creates log_welfare and converts categorical variables into factors.
# Build the stacked frame mice will work on: GLSS train rows (.role "train",
# welfare observed) on top of the apply rows (.role "apply", welfare masked).
# Stacking is what lets one model be FIT on GLSS and PREDICTED into the apply
# rows (slide "Synthetic data is a missing data framework").
build_frame <- function(apply_dat, targets_to_mask, extra_cols = character(0)) {
  cols <- c("hh_key", "household_weight", core_pred,
            "welfare", "nhis_any_member_registered", extra_cols)
  tr <- glss_train %>% select(any_of(cols)) %>% mutate(.role = "train")
  ap <- apply_dat %>% select(any_of(cols)) %>% mutate(.role = "apply")
  stacked <- bind_rows(tr, ap) %>%
    mutate(log_welfare = ifelse(!is.na(welfare) & welfare > 0, log(welfare), NA_real_))
  # Hide the targets on the apply rows so mice treats them as "to be imputed".
  for (v in targets_to_mask) stacked[[v]][stacked$.role == "apply"] <- NA
  for (v in intersect(c(cat_predictors, "head_educ_cat", "toilet_type_screen"),
                      names(stacked))) stacked[[v]] <- factor(stacked[[v]])
  stacked$region10 <- factor(stacked$region10)
  stacked$nhis_any_member_registered <-
    factor(stacked$nhis_any_member_registered, levels = c(0, 1))
  if ("has_electricity" %in% names(stacked))
    stacked$has_electricity <- factor(stacked$has_electricity, levels = c(0, 1))
  stacked
}

# Fit mice and draw m imputed datasets. The three decisions from the slide
# "Decisions for every imputed variable" appear here:
#   (1) target_methods   = the ENGINE per target (norm for welfare, logreg for binary)
#   (2) target_predictors = the CONDITIONING set X for each target
#   (3) ignore = .role=="apply" = fit PARAMETERS on GLSS train rows only; apply
#       rows receive draws but never inform the model (slide "Step 1: Choose
#       a data set to build the model on").
run_mice <- function(frame, target_methods, target_predictors) {
  mice_cols <- setdiff(names(frame), c("hh_key", "household_weight",
                                       "welfare", ".role", collinear_drop))
  mdat <- frame[, mice_cols]
  meth <- make.method(mdat)
  # make.method() also assigns default methods to any shared predictors X that
  # have incidental missing cells. In these prepared teaching files this is a
  # tiny amount of missingness, so mice may fill those X cells while the
  # substantive target remains the missing welfare/NHIS/electricity block.
  for (v in names(target_methods)) meth[v] <- target_methods[v]
  pm <- make.predictorMatrix(mdat)
  for (v in names(target_predictors)) {
    pm[v, ] <- 0
    pm[v, intersect(target_predictors[[v]], colnames(pm))] <- 1
  }
  mice(mdat, m = M, maxit = MAXIT, seed = SYNTH_SEED,
       method = meth, predictorMatrix = pm,
       ignore = frame$.role == "apply", printFlag = FALSE)
}

# This extracts the completed/imputed values from the fitted mice object.
# Pull the apply-row draws out of a fitted mice object, one stacked row per
# (household, imputed dataset). `synth_*` columns hold the IMPUTED values; poverty is
# DERIVED from each drawn welfare here, not drawn.
collect_draws <- function(imp, frame) {
  bind_rows(lapply(seq_len(M), function(i) {
    comp <- complete(imp, i)
    frame %>% select(hh_key, household_weight, .role) %>%
      mutate(.imp = i,
             synth_log_welfare = comp$log_welfare,
             synth_welfare = exp(comp$log_welfare),          # back-transform log -> GHS
             synth_nhis = as.integer(as.character(comp$nhis_any_member_registered)),
             synth_poverty_status = as.character(derive_poverty(exp(comp$log_welfare))),
             synth_poor = as.integer(exp(comp$log_welfare) < POV_ABSOLUTE)) %>%
      filter(.role == "apply") %>% select(-.role)            # keep only the imputed rows
  }))
}

## RUN A: DHS enrichment ---------------------------------------------------
# The real goal: fit on the SAME GLSS train rows, then impute welfare + NHIS into
# the DHS households (whose welfare is missing by design). Each DHS household
# leaves with m drawn welfare values and m derived poverty labels (slides
# "The imputation workflow at a glance" / "Step 2: partially synthetic data").
frame_a <- build_frame(dhs_households, c("log_welfare", "nhis_any_member_registered"))
imp_a <- run_mice(frame_a,
  c(log_welfare = "norm", nhis_any_member_registered = "logreg"),
  list(log_welfare = c(model_pred, "nhis_any_member_registered"),
       nhis_any_member_registered = c(model_pred, "log_welfare")))
draws_a <- collect_draws(imp_a, frame_a)

# This joins each DHS household's imputed welfare/NHIS draws to observed DHS
# descriptors we use for summaries and benchmark comparisons.
dhs_synth <- draws_a %>%
  left_join(dhs_households %>% select(hh_key, region10, urban, wealth_index, wealth_score,
                                      nhis_observed = nhis_any_member_registered),
            by = "hh_key")

# This summarizes the DHS enrichment separately for each imputed dataset.
enrichment_by_imp <- dhs_synth %>% group_by(.imp) %>%
  summarise(synth_welfare_mean_w = round(w_mean(synth_welfare, household_weight), 1),
            synth_welfare_p50 = round(median(synth_welfare, na.rm = TRUE), 1),
            synth_poverty_rate_w = round(100 * w_mean(synth_poor, household_weight), 2),
            synth_nhis_rate_w = round(100 * w_mean(synth_nhis, household_weight), 2),
            observed_nhis_rate_w = round(100 * w_mean(nhis_observed, household_weight), 2),
            .groups = "drop")

# This attaches household-level synthetic poverty to each DHS child record so
# child outcomes can be analyzed by the imputed household poverty status.
dhs_children_synth <- dhs_children %>%
  inner_join(dhs_synth %>% select(hh_key, .imp, synth_poor, synth_poverty_status),
             by = "hh_key", relationship = "many-to-many")

# This estimates the stunting gradient by the three synthetic poverty-status
# groups across imputed datasets, keeping between-imputation variation visible.
gradient <- dhs_children_synth %>%
  group_by(.imp, synth_poverty_status) %>%
  summarise(stunted_rate = 100 * w_mean(stunted, child_weight), .groups = "drop") %>%
  group_by(synth_poverty_status) %>%
  summarise(stunted_rate_pooled = round(mean(stunted_rate), 2),
            between_imp_sd = round(sd(stunted_rate), 2), .groups = "drop")

# This collects the main DHS enrichment summaries for later tables/output.
enrichment_summary <- bind_rows(
  enrichment_by_imp %>% mutate(table = "enrichment_by_imputed_dataset"),
  gradient %>% mutate(table = "stunting_by_synth_poverty_status"))

# This creates a DHS-only wealth-index benchmark for comparison with the
# synthetic poverty gradient.
wealth_benchmark <- dhs_children %>%
  inner_join(dhs_households %>% select(hh_key, wealth_index), by = "hh_key") %>%
  mutate(low_wealth = as.integer(wealth_index <= 2)) %>%
  group_by(low_wealth) %>%
  summarise(stunted_rate = round(100 * w_mean(stunted, child_weight), 2), .groups = "drop")

dhs_overall_stunting <- data.frame(
  group = "All DHS children",
  stunted_rate = round(w_rate(dhs_children$stunted, dhs_children$child_weight), 2),
  between_imp_sd = NA_real_
)

## RUN B: GLSS holdout (Validation Check 1) --------------------------------
# Fit on GLSS train, impute welfare + NHIS into the held-out GLSS rows whose
# TRUE welfare we have hidden. This is the only place we can score imputed
# welfare against ground truth (slide "Validation Check 1 — GLSS holdout").
# Note the chained factorisation (slide "Activity: factor the model"): each
# target conditions on the other (welfare | X, NHIS) and (NHIS | X, welfare).
frame_b <- build_frame(glss_holdout, c("log_welfare", "nhis_any_member_registered"))
imp_b <- run_mice(frame_b,
  c(log_welfare = "norm", nhis_any_member_registered = "logreg"),
  list(log_welfare = c(model_pred, "nhis_any_member_registered"),
       nhis_any_member_registered = c(model_pred, "log_welfare")))
draws_b <- collect_draws(imp_b, frame_b)

# Recover the TRUE holdout welfare we hid, then join imputed-to-true row by row.
# This rebuilds the “answer key” for the GLSS holdout sample.
truth_b <- glss_holdout %>%
  transmute(hh_key, true_welfare = welfare,
            true_poor = as.integer(welfare < POV_ABSOLUTE),
            true_status = as.character(derive_poverty(welfare)),
            true_nhis = nhis_any_member_registered,
            region10, urban, household_weight)

# This joins the synthetic/imputed values to the true holdout values household by household.
val_b <- draws_b %>% inner_join(truth_b, by = "hh_key") %>%
  rename(household_weight = household_weight.x) %>% select(-household_weight.y)
holdout_synth <- val_b

# This calculates validation statistics separately for each imputed dataset.
per_imp <- val_b %>% group_by(.imp) %>%
  summarise(
    synth_welfare_mean_w = w_mean(synth_welfare, household_weight),
    true_welfare_mean_w = w_mean(true_welfare, household_weight),
    synth_poverty_rate_w = 100 * w_mean(synth_poor, household_weight),
    true_poverty_rate_w = 100 * w_mean(true_poor, household_weight),
    mae_log = mean(abs(synth_log_welfare - log(true_welfare)), na.rm = TRUE),
    rmse_log = sqrt(mean((synth_log_welfare - log(true_welfare))^2, na.rm = TRUE)),
    corr_log = cor(synth_log_welfare, log(true_welfare), use = "complete.obs"),
    spearman = cor(synth_welfare, true_welfare, method = "spearman", use = "complete.obs"),
    poverty3_accuracy = mean(synth_poverty_status == true_status, na.rm = TRUE),
    nhis_synth_rate_w = 100 * w_mean(synth_nhis, household_weight),
    nhis_true_rate_w = 100 * w_mean(true_nhis, household_weight),
    nhis_accuracy = mean(synth_nhis == true_nhis, na.rm = TRUE), .groups = "drop")

# This formats the per_imp results into a table for teaching/reporting.
holdout_validation <- bind_rows(
  per_imp %>% mutate(.imp = as.character(.imp)),
  per_imp %>% summarise(across(-.imp, mean)) %>% mutate(.imp = "pooled_mean"),
  per_imp %>% summarise(across(-.imp, sd)) %>% mutate(.imp = "between_imp_sd")
) %>% mutate(across(where(is.numeric), ~ round(.x, 4))) %>% select(.imp, everything())

# This creates a confusion table for poverty status.
holdout_confusion <- val_b %>% count(true_status, synth_poverty_status) %>%
  mutate(n = n / M) %>% arrange(true_status, synth_poverty_status)

# This is a small helper for subgroup validation
# within each imputed dataset and within each group, compare synthetic poverty rates to true poverty rates.
subgrp <- function(dat, gvar) {
  dat %>% group_by(.imp, grp = .data[[gvar]]) %>%
    summarise(synth_pov_w = 100 * w_mean(synth_poor, household_weight),
              true_pov_w = 100 * w_mean(true_poor, household_weight), .groups = "drop") %>%
    group_by(grp) %>%
    summarise(group_var = gvar,
              synth_poverty_rate_w = round(mean(synth_pov_w), 2),
              true_poverty_rate_w = round(mean(true_pov_w), 2),
              gap_pp = round(mean(synth_pov_w - true_pov_w), 2), .groups = "drop")
}
holdout_subgroups <- bind_rows(subgrp(val_b, "region10"), subgrp(val_b, "urban"))

## RUN C: DHS electricity transport check (Validation Check 2) --------------
# The "calibration" check from the literature (slide "Validation Check 2 — transport").
# Electricity is a SHARED variable DHS actually observes. We pretend DHS lacks
# it, predict it from the GLSS-trained model, then compare to the truth DHS
# recorded. It tests whether a GLSS relationship transports to DHS -- on an
# observed target -- WITHOUT needing welfare ground truth.
frame_c <- build_frame(dhs_households, "has_electricity", extra_cols = "has_electricity")
frame_c$log_welfare <- NULL; frame_c$welfare <- NULL
frame_c$nhis_any_member_registered <- NULL
imp_c <- run_mice(frame_c, c(has_electricity = "logreg"),
                  list(has_electricity = model_pred))

# This pulls the imputed DHS electricity draws and joins the true DHS
# electricity value we temporarily hid.
calib <- bind_rows(lapply(seq_len(M), function(i) {
  comp <- complete(imp_c, i)
  frame_c %>% select(hh_key, household_weight, .role, region10) %>%
    mutate(.imp = i, synth_elec = as.integer(as.character(comp$has_electricity))) %>%
    filter(.role == "apply")
})) %>%
  left_join(dhs_households %>% select(hh_key, true_elec = has_electricity, urban), by = "hh_key")

# This calculates the transport-check performance separately for each imputed dataset.
calib_per_imp <- calib %>% group_by(.imp) %>%
  summarise(synth_elec_rate_w = round(100 * w_mean(synth_elec, household_weight), 2),
            true_elec_rate_w = round(100 * w_mean(true_elec, household_weight), 2),
            accuracy = round(mean(synth_elec == true_elec, na.rm = TRUE), 4), .groups = "drop")

# This compares synthetic and true electricity rates within policy-relevant subgroups.
calib_subgrp <- bind_rows(
  calib %>% group_by(grp = as.character(region10)) %>%
    summarise(group_var = "region10",
              synth_rate_w = round(100 * w_mean(synth_elec, household_weight), 2),
              true_rate_w = round(100 * w_mean(true_elec, household_weight), 2), .groups = "drop"),
  calib %>% group_by(grp = as.character(urban)) %>%
    summarise(group_var = "urban",
              synth_rate_w = round(100 * w_mean(synth_elec, household_weight), 2),
              true_rate_w = round(100 * w_mean(true_elec, household_weight), 2), .groups = "drop")
) %>% mutate(gap_pp = round(synth_rate_w - true_rate_w, 2))

# This collects the electricity transport-check summaries for later tables/output.
electricity_calibration <- bind_rows(
  calib_per_imp %>% mutate(table = "by_imputed_dataset", grp = NA, group_var = NA,
                           synth_rate_w = synth_elec_rate_w, true_rate_w = true_elec_rate_w) %>%
    select(table, .imp, group_var, grp, synth_rate_w, true_rate_w, accuracy),
  calib_subgrp %>% mutate(table = "subgroup", .imp = NA, accuracy = NA) %>%
    select(table, .imp, group_var, grp, synth_rate_w, true_rate_w, accuracy, gap_pp))

## model_spec + logged_events ----------------------------------------------
model_spec <- data.frame(
  item = c("data_family", "mice_version", "seed", "m", "maxit",
           "fit_rows_rule", "weights_statement", "welfare_method",
           "welfare_transform", "nhis_method", "electricity_method",
           "poverty_rule", "primary_predictors", "collinear_dropped_from_model",
           "transport_check_predictors", "sensitivity_predictors_used"),
  value = c(DATA_FAMILY, as.character(packageVersion("mice")), SYNTH_SEED, M, MAXIT,
    "mice(ignore=): models fit on GLSS train rows only; holdout/DHS rows receive draws without informing parameters",
    "UNWEIGHTED fitting (mice norm/logreg take no case weights); household_weight used in all summaries",
    "norm (Bayesian linear regression; parameter draws then value draws)",
    "log(welfare); draws back-transformed exp(); positive by construction",
    "logreg (active_sensitivity target; masked in DHS though observed, so the DHS comparison is also a transport diagnostic)",
    "logreg (has_electricity excluded from its own conditioning set; welfare/NHIS not in the transport-check frame)",
    sprintf("derived from drawn welfare at %.2f / %.0f; never drawn directly", POV_EXTREME, POV_ABSOLUTE),
    paste(model_pred, collapse = "; "),
    "members_total, adults_15plus, floor_non_earth: exact linear dependencies; dropped from the MODEL set only, kept in extract/manifest",
    paste(model_pred, collapse = "; "), "none in the primary specification"),
  stringsAsFactors = FALSE)

primary_predictors <- model_spec %>%
  filter(item == "primary_predictors") %>%
  pull(value) %>%
  strsplit("; ", fixed = TRUE) %>%
  unlist()

logged_events <- bind_rows(lapply(
  list(A_enrichment = imp_a, B_holdout = imp_b, C_calibration = imp_c),
  function(imp) {
    le <- imp$loggedEvents
    if (is.null(le)) return(data.frame(n_events = 0L, detail = "none"))
    le %>% count(dep, meth, out, name = "n_events") %>%
      transmute(n_events, detail = paste0(dep, " | ", meth, " | ", out))
  }) %>% setNames(c("A_enrichment", "B_holdout", "C_calibration")), .id = "run")
```

# Child Stunting Outcomes by Imputed Poverty in DHS

*Theory slide: "Running question for today" — stunting by consumption-based poverty status.*

The imputation chunk has already transported GLSS-style welfare and NHIS into DHS households. We now use those household-level imputed variables to answer the running DHS child-stunting question.

## Transporting Imputed Variables from GLSS to DHS

*Theory slides: "Step 2: We create partially synthetic data" and "From the welfare aggregate to poverty".*

Welfare is drawn on the log scale and back-transformed. Poverty status is derived from each drawn welfare value at the official GLSS6 lines; it is never drawn as its own outcome.

```{r enrichment-summary-table}
enrichment_summary %>%
  filter(table == "enrichment_by_imputed_dataset") %>%
  select(.imp, synth_welfare_mean_w, synth_welfare_p50,
         synth_poverty_rate_w, synth_nhis_rate_w, observed_nhis_rate_w) %>%
  knitr::kable()
```

How to read this enrichment table: poverty and welfare should be stable across imputed datasets. The NHIS synthetic-vs-observed gap is not a welfare validation result; it is a reminder that even observed DHS targets can show transport gaps.

```{r enrichment-poverty-plot}
enrichment_summary %>%
  filter(table == "enrichment_by_imputed_dataset") %>%
  ggplot(aes(x = factor(.imp), y = synth_poverty_rate_w)) +
  geom_col(fill = "#2f6f73") +
  labs(x = "Imputed dataset", y = "Weighted synthetic poverty rate (%)") +
  theme_minimal(base_size = 12)
```

How to read this enrichment figure: the bars should be close to one another. Big bar-to-bar movement would mean the substantive child-outcome comparison depends too much on the random imputed dataset.

```{r derive-not-draw-check}
# Proof that poverty is DERIVED, not drawn (slide "From the welfare aggregate
# to poverty"). We re-cut the imputed welfare here and confirm it equals
# the poverty label stored during collect_draws(): mismatches should be zero.
derive_check <- dhs_synth %>%
  mutate(
    derived_status = as.character(cut(
      synth_welfare,
      c(-Inf, POV_EXTREME, POV_ABSOLUTE, Inf),
      labels = c("Very poor", "Poor", "Non poor"))),
    derived_poor = as.integer(synth_welfare < POV_ABSOLUTE)
  ) %>%
  summarise(
    rows = n(),
    status_mismatches = sum(derived_status != synth_poverty_status, na.rm = TRUE),
    poor_mismatches = sum(derived_poor != synth_poor, na.rm = TRUE)
  )

knitr::kable(derive_check)
```

How to read this check: zero mismatches is the target. Poverty status is a deterministic consequence of each synthetic welfare draw, not a separately modeled category.

## Overall and by imputed poverty status

DHS child records are linked to each household imputed dataset. Each child therefore appears once per imputed dataset, carrying the same observed child outcome but a potentially different imputed household poverty status.

```{r child-linkage}
child_link_summary <- data.frame(
  rows = nrow(dhs_children_synth),
  imputed_datasets = length(unique(dhs_children_synth$.imp)),
  child_rows_per_imputed_dataset = nrow(dhs_children_synth) / length(unique(dhs_children_synth$.imp)),
  unique_child_ids = length(unique(dhs_children_synth$child_id))
)

knitr::kable(child_link_summary)
```

```{r stunting-gradient}
poverty_levels <- c("Very poor", "Poor", "Non poor")

stunting_by_imp <- dhs_children_synth %>%
  mutate(synth_poverty_status = factor(synth_poverty_status, levels = poverty_levels)) %>%
  group_by(.imp, synth_poverty_status) %>%
  summarise(stunted_rate = 100 * w_mean(stunted, child_weight), .groups = "drop")

stunting_pooled <- stunting_by_imp %>%
  group_by(synth_poverty_status) %>%
  summarise(between_imp_sd = sd(stunted_rate),
            stunted_rate = mean(stunted_rate), .groups = "drop") %>%
  mutate(group = paste("Synthetic", tolower(as.character(synth_poverty_status))),
         group = gsub("non poor", "non-poor", group, fixed = TRUE),
         across(c(stunted_rate, between_imp_sd), ~ round(.x, 2))) %>%
  select(group, stunted_rate, between_imp_sd)

stunting_display <- bind_rows(dhs_overall_stunting, stunting_pooled)

knitr::kable(stunting_display)
```

How to read this gradient: start with the overall DHS stunting rate, then compare the three imputed poverty groups. The between-imputation SD tells us how much each group rate moves across synthetic welfare draws.

```{r wealth-benchmark}
wealth_benchmark %>%
  mutate(group = ifelse(low_wealth == 1,
                        "DHS wealth bottom two quintiles",
                        "DHS wealth upper three quintiles")) %>%
  select(group, stunted_rate) %>%
  knitr::kable()
```

```{r child-gradient-plot}
overall_stunting_rate <- dhs_overall_stunting$stunted_rate[1]

stunting_pooled %>%
  ggplot(aes(x = group, y = stunted_rate, fill = group)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_hline(yintercept = overall_stunting_rate, linetype = "dashed",
             color = "#262626") +
  annotate("text", x = 2.6, y = overall_stunting_rate + 1,
           label = "Overall DHS rate", size = 3.5, hjust = 0) +
  labs(x = NULL, y = "Weighted stunting rate (%)") +
  theme_minimal(base_size = 12)
```

# Pooling Across Imputed Datasets

Multiple imputed datasets keep synthetic-data uncertainty visible. For a scalar estimate, combine the average within-imputation uncertainty and the between-imputation variation.

The small calculation below collapses very poor and poor households into "synthetic poor" and pools the poor-minus-nonpoor stunting-rate gap. The within-imputation variance uses a simple effective-sample-size approximation for weighted binary means, which is fine for a classroom diagnostic. A publication analysis would use a survey-design-aware variance estimator.

```{r rubin-style-pooling}
# Step 1: compute the SAME estimate on each imputed dataset separately -- here
# the poor-minus-nonpoor stunting gap, plus its within-imputation variance.
gap_by_imp <- bind_rows(lapply(split(dhs_children_synth, dhs_children_synth$.imp), function(dat) {
  poor <- dat %>% filter(synth_poor == 1)
  nonpoor <- dat %>% filter(synth_poor == 0)
  poor_est <- weighted_binary_est(poor$stunted, poor$child_weight)
  nonpoor_est <- weighted_binary_est(nonpoor$stunted, nonpoor$child_weight)
  data.frame(.imp = unique(dat$.imp),
             gap_pp = 100 * (poor_est$estimate - nonpoor_est$estimate),
             within_var_pp2 = 100^2 * (poor_est$var + nonpoor_est$var))
}))

# Step 2: combine the m estimates with Rubin's rules. The reported SE now
# carries BOTH the usual sampling noise AND the extra spread from imputing
# welfare (slide "Multiple imputation keeps uncertainty visible").
pooled_gap <- pool_scalar(gap_by_imp$gap_pp, gap_by_imp$within_var_pp2) %>%
  mutate(across(everything(), ~ round(.x, 3)))

knitr::kable(gap_by_imp %>% mutate(across(where(is.numeric), ~ round(.x, 3))))
knitr::kable(pooled_gap)
```

# Holdout Check: Can We Recover GLSS Welfare?

*Theory slide: "Validation Check 1 — GLSS holdout".*

Before enriching DHS, hide welfare in the GLSS holdout and impute it as if it were missing. This is the only check where we observe true GLSS welfare for the rows receiving imputed welfare, so it is where we can actually score the model against ground truth.

```{r holdout-summary}
# One row per imputed dataset plus pooled mean and between-imputation SD.
# Read across a row: imputed poverty rate should track the true rate.
knitr::kable(holdout_validation)
```

How to read this table: the synthetic poverty rate should be close to the true holdout rate, and the per-imputation rows should not swing wildly. The log-welfare correlation is not expected to be near one because housing and demographic predictors do not fully explain household welfare, and posterior-predictive draws deliberately add uncertainty.

```{r holdout-poverty-plot}
holdout_plot <- holdout_validation %>%
  filter(.imp != "between_imp_sd") %>%
  mutate(.imp = factor(.imp, levels = c(as.character(seq_len(M)), "pooled_mean"))) %>%
  select(.imp, synth_poverty_rate_w, true_poverty_rate_w) %>%
  tidyr::pivot_longer(-.imp, names_to = "series", values_to = "poverty_rate")

ggplot(holdout_plot, aes(x = .imp, y = poverty_rate, color = series, group = series)) +
  geom_point(size = 2.2) +
  geom_line() +
  labs(x = "Imputed dataset", y = "Weighted poverty rate (%)", color = NULL) +
  theme_minimal(base_size = 12)
```

How to read this figure: a useful model keeps the synthetic line near the true line and keeps the imputed datasets stable. A large spread across imputed datasets would mean the imputation uncertainty is dominating the lesson.

```{r holdout-density-overlay}
truth_density <- holdout_synth %>%
  distinct(hh_key, true_welfare) %>%
  mutate(log_welfare = log(true_welfare), series = "True holdout welfare")

synth_density <- holdout_synth %>%
  mutate(log_welfare = synth_log_welfare,
         series = paste("Synthetic dataset", .imp))

ggplot() +
  geom_density(data = synth_density, aes(x = log_welfare, group = .imp),
               color = "#2f6f73", alpha = 0.28, linewidth = 0.8) +
  geom_density(data = truth_density, aes(x = log_welfare),
               color = "#262626", linewidth = 1.2) +
  labs(x = "Log welfare", y = "Density") +
  theme_minimal(base_size = 12)
```

How to read this density check: the synthetic distributions should cover the same main range and shape as the true holdout distribution. Tail misses are common; a large shift in the whole density would be a warning.

```{r holdout-confusion}
holdout_confusion %>%
  tidyr::pivot_wider(names_from = synth_poverty_status,
                     values_from = n,
                     values_fill = 0) %>%
  knitr::kable()
```

How to read this confusion matrix: most mass should be on the diagonal. Adjacent-category confusion is expected, especially near the poverty line; opposite-corner mass is the alarm signal.

```{r holdout-subgroups}
holdout_subgroups %>%
  mutate(across(c(synth_poverty_rate_w, true_poverty_rate_w, gap_pp),
                ~ round(.x, 1))) %>%
  arrange(group_var, grp) %>%
  knitr::kable()
```

How to read this subgroup check: national fit is not enough. Good synthetic data should preserve the poverty pattern across policy-relevant groups; large regional or urban/rural gaps identify where the transport story is weakest.

The holdout result is not perfect, but it is useful. The pooled log-welfare correlation is about `r round(holdout_validation$corr_log[holdout_validation$.imp == "pooled_mean"], 2)`: person-level welfare prediction is bounded by what housing and demographic variables can explain, and posterior-predictive draws add deliberate noise. The success criteria here are distributional recovery, subgroup diagnostics, and policy-gradient recovery, not one-for-one household prediction.

# DHS Transport Check: How well does imputed electricity transfer to DHS

*Theory slide: "Validation Check 2 — transport" (the literature calls this **calibration**).*

The transport check uses an observed DHS variable as a **validation target**. Here the model hides DHS electricity, predicts it from GLSS-trained relationships, and compares predicted electricity with true DHS electricity. This checks transport for an observed target, not welfare itself — so it works even though DHS never observes welfare.

```{r transport-check}
electricity_calibration %>%
  filter(table == "by_imputed_dataset") %>%
  select(.imp, synth_rate_w, true_rate_w, accuracy) %>%
  knitr::kable()
```

How to read this transport check: stable predicted rates across imputed datasets mean parameter uncertainty is modest. A persistent level gap means the GLSS-trained relationship does not fully transport to DHS for this observed target.

```{r transport-subgroups}
electricity_calibration %>%
  filter(table == "subgroup", group_var %in% c("region10", "urban")) %>%
  mutate(gap_pp = round(synth_rate_w - true_rate_w, 1)) %>%
  arrange(group_var, grp) %>%
  knitr::kable()
```

How to read this subgroup transport check: the national gap can hide where the model is doing well or poorly. Region and urban/rural gaps are the evidence participants should carry into the real welfare enrichment.

The level gap is a teaching feature. It says the model trained on 2012/13 GLSS relationships under-predicts a 2014 DHS household target in some groups. That is the right warning to carry into the real welfare enrichment.

# Appendix: Optional Engine Sensitivity

The primary pipeline uses `norm` for log welfare because it is transparent and directly matches the posterior-predictive story in the theory slides. This optional appendix lets you switch the welfare engine on a small holdout sample. It is set to `eval = FALSE` so it does not run during the main practical; run it manually, or change the chunk option to `eval = TRUE`, when you want the sensitivity exercise.

```{r extension-engine-switch, eval=FALSE}
engine <- "pmm"
allowed_engines <- c("norm", "pmm", "cart", "rf")
if (!engine %in% allowed_engines) {
  stop("Choose one of: ", paste(allowed_engines, collapse = ", "))
}

run_engine_holdout <- function(engine, n_train = 1500, n_holdout = 500,
                               m = 2, maxit = 2) {
  set.seed(EXTENSION_SEED)
  use_cols <- c("hh_key", "household_weight", primary_predictors, "welfare")

  tr_pool <- glss_train %>%
    select(all_of(use_cols)) %>%
    filter(if_all(all_of(c(primary_predictors, "welfare")), ~ !is.na(.x))) %>%
    arrange(hh_key)
  ho_pool <- glss_holdout %>%
    select(all_of(use_cols)) %>%
    filter(if_all(all_of(c(primary_predictors, "welfare")), ~ !is.na(.x))) %>%
    arrange(hh_key)

  tr <- tr_pool %>% slice_sample(n = min(n_train, nrow(tr_pool))) %>% mutate(.role = "train")
  ho <- ho_pool %>% slice_sample(n = min(n_holdout, nrow(ho_pool))) %>% mutate(.role = "holdout")

  stacked <- bind_rows(tr, ho) %>%
    mutate(log_welfare = ifelse(.role == "train", log(welfare), NA_real_))
  for (v in intersect(primary_predictors, names(stacked)))
    if (is.character(stacked[[v]])) stacked[[v]] <- factor(stacked[[v]])
  if ("region10" %in% names(stacked)) stacked$region10 <- factor(stacked$region10)

  mdat <- stacked %>% select(all_of(c(primary_predictors, "log_welfare")))
  meth <- mice::make.method(mdat); meth[] <- ""; meth["log_welfare"] <- engine
  pm <- mice::make.predictorMatrix(mdat); pm[, ] <- 0
  pm["log_welfare", primary_predictors] <- 1

  elapsed <- system.time({
    imp <- mice::mice(mdat, m = m, maxit = maxit, method = meth, predictorMatrix = pm,
                      ignore = stacked$.role == "holdout", seed = EXTENSION_SEED, printFlag = FALSE)
  })

  draws <- bind_rows(lapply(seq_len(m), function(i) {
    comp <- mice::complete(imp, i)
    holdout_idx <- stacked$.role == "holdout"
    stacked %>% filter(.role == "holdout") %>%
      transmute(hh_key, household_weight, .imp = i,
                synth_log_welfare = comp$log_welfare[holdout_idx],
                synth_welfare = exp(synth_log_welfare),
                synth_poor = as.integer(synth_welfare < POV_ABSOLUTE),
                true_welfare = welfare, true_poor = as.integer(welfare < POV_ABSOLUTE))
  }))

  recovery_metrics(draws, engine) %>%
    mutate(train_rows = nrow(tr), holdout_rows = nrow(ho),
           runtime_sec = round(elapsed[["elapsed"]], 1))
}

extension_results <- bind_rows(
  run_engine_holdout("norm"),
  if (engine == "norm") NULL else run_engine_holdout(engine)
) %>%
  mutate(across(c(mae_log, rmse_log, corr_log,
                  synth_poverty_rate_w, true_poverty_rate_w), ~ round(.x, 3)))

knitr::kable(extension_results)
```

How to read this extension: lower MAE/RMSE and higher correlation are better household-level recovery, but the poverty-rate comparison still matters because the policy analysis uses poverty groups. Tree and forest engines can tighten nonlinear fit, but they can also change tails and subgroup behavior.

## Optional self-guided exercises

The optional appendix chunk above is designed so you can change one thing at a time and re-run manually.

1. Change `engine` to `"pmm"`, `"cart"`, or `"rf"` and compare the GLSS holdout metrics against `"norm"`.
2. Remove one weak predictor from the model and check whether the holdout or transport check changes.
3. Treat NHIS as a sensitivity target: does the model reproduce the DHS level gap or erase it?
4. Focus on one subgroup, such as rural households or a region, and decide whether the enrichment is credible enough for that policy question.

The key question is not whether one number changes. The key question is whether the holdout, the transport check, and the policy gradient tell a coherent story.

# Appendix: Predictor and Target Roles

The practical uses a manifest built from the bridge screening step. This keeps the notebook synchronized with the data-preparation decisions.

```{r manifest-roles}
role_counts <- manifest %>%
  count(manifest_role, sort = TRUE)

knitr::kable(role_counts)
```

```{r model-spec}
model_table <- data.frame(
  design_choice = c("Fitting rows", "Weights", "Welfare target",
                    "NHIS target", "Poverty status", "Primary predictors"),
  implementation = c(
    model_spec$value[model_spec$item == "fit_rows_rule"],
    model_spec$value[model_spec$item == "weights_statement"],
    paste(model_spec$value[model_spec$item == "welfare_method"],
          model_spec$value[model_spec$item == "welfare_transform"], sep = "; "),
    model_spec$value[model_spec$item == "nhis_method"],
    model_spec$value[model_spec$item == "poverty_rule"],
    paste(primary_predictors, collapse = ", ")
  )
)

knitr::kable(model_table)
```

Three core-manifest columns are deliberately dropped from the model: `members_total` and `adults_15plus` are exact linear combinations of other household-composition predictors, while `floor_non_earth` duplicates information already carried by `floor_material_screen`. The columns remain in the extract for inspection.

```{r model-drops}
dropped <- model_spec %>%
  filter(item == "collinear_dropped_from_model") %>%
  pull(value)

dropped
```
